---
title: "World suicide analysis"
author: Eveline Xu xuehui
date: 2020/12/1

output: html_notebook
---

```{r Outline}
```
# Table of contents:
1) Data source
2) Preprocessing
3) Data Analysis
4) Overall view and Foundings
5) Interpretation and Suggestion
```{r}
```
# Data source
Suicide data. (2020). from https://www.who.int/teams/mental-health-and-substance-use/suicide-data

# References
- World Health Organization. (2018). Suicide prevention，from http://www.who.int/mental_health/suicide-prevention/en/

- "Gender Differences In Suicide". 2020. En.Wikipedia.Org. https://en.wikipedia.org/wiki/Gender_differences_in_suicide#:~:text=Globally%2C%20death%20by%20suicide%20occurred,over%20the%20age%20of%2065.

```{r}
```
### Importing Data and Libraries
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
suicideData <- read_csv("/Users/glameveline/RProjects/Fianl project/suicideData.csv")
head(suicideData)
```
```{r md}
```
# Preprocessing
### Dealling with missing values
```{r}
summary(suicideData)
```
Found missing values in suicides_no and population
```{r}
table(is.na(suicideData$suicides_no))
table(is.na(suicideData$population))
```
## Data analysis
#### - Rate of suicide per 100k in 1995, 2005 and 2015
```{r by year line}
suicide_year <- suicideData %>%
  filter(!is.na(population) & !is.na(suicides_no)) %>%
  select(year,suicides_no,population) %>%
  group_by(year) %>%
  summarise(suicide_ratio=round((sum(suicides_no)/sum(population))*100000,2))
suicide_year
```
```{r vidualization1}
ggplot(data=suicide_year,aes(x=year,y=suicide_ratio))+geom_line()+
  geom_point(colour = "red") +
  geom_text(aes(label = suicide_ratio, vjust =-2, hjust =0.5))+
  ylim(0,20)+
  labs(
    title = "World suicide rate trend over 20 years",
    subtitle = "1995,2005 and 2005",
    y = "Suicide Rate per 100K")
```

#### - suicide rate by age group from 1995 to 2015

```{r}
suicide_age <-  suicideData %>%
  filter(!is.na(population) & !is.na(suicides_no)) %>%
  select(year,age,suicides_no,population) %>%
  group_by(age,year) %>%
  summarise(suicide_ratio=round((sum(suicides_no)/sum(population))*100000,2))

suicide_age


ggplot(data=suicide_age,aes(x=year,y=suicide_ratio, group=age,color=age))+
  geom_line()+
  geom_point()+
        geom_text(aes(label = suicide_ratio, vjust =-2, hjust =0.5))+
  ggtitle("World suicide rate by age group in the previous 20 years")+
  labs(
    y="Suicide Rate per 100K"
  )
```

#### - Top 10 suicide rate regions in 2015
```{r}
# top 10 per 100 k by region(country) in 2015
top10_suicide_country <- suicideData %>%
  filter(!is.na(population) & !is.na(suicides_no) & year=='2015') %>%
  select(country,suicides_no,population) %>%
  group_by(country) %>%
  summarise(suicide_ratio=round((sum(suicides_no)/sum(population))*100000,2)) %>%
  arrange(desc(suicide_ratio)) %>%
  head(10)

top10_suicide_country


plot_ly(data = top10_suicide_country,x=~reorder(country,-suicide_ratio),y=~suicide_ratio,type ='bar',color=top10_suicide_country$country,colors=c('Paired')) %>%
  layout(title='World Top10 suicide commited regions in 2015',
         yaxis=list(title='Suicide rate per 100K'),xaxis=list(title='Country'))
```
#### - Choropleth map of suicide rate in the year of 2015
```{r}
# Choropleth map of suicide rate in the year of 2015

suicide_country <- suicideData %>%
  filter(!is.na(population) & !is.na(suicides_no) & year=='2015') %>%
  select(country,suicides_no,population) %>%
  group_by(country) %>%
  summarise(suicide_ratio=round((sum(suicides_no)/sum(population))*100000,2))
suicide_country <- rename(suicide_country,region=country)

# importing world map
world_map <- map_data('world')
str(world_map)
# changing names of certain countries
world_map$region <- ifelse(world_map$region=='USA',"United States of America",world_map$region)
head(world_map,n=20)

# suicide map
world_suicide <- left_join(world_map,suicide_country,by='region')
head(world_suicide)

# vidualization
arrange(world_suicide, order) %>%
  ggplot(aes(long, lat, group = group, fill =suicide_ratio)) +
  geom_polygon() +
  xlim(-110,150)+ylim(-20,90)+
  coord_map("polyconic") +
  theme(panel.grid =element_blank())+
  scale_fill_continuous(low = "green", high = "red")+
  labs(
    y='',
    x=''
  )
```
#### - World suicide rate per 100K by gender
```{r}
suicide_gender <- suicideData %>%
  filter(!is.na(population) & !is.na(suicides_no)) %>%
  select(sex,year,suicides_no,population) %>%
  group_by(sex,year) %>%
  summarise(suicide_ratio=round((sum(suicides_no)/sum(population))*100000,2))
suicide_gender

# vidualization
plot_ly(data = suicide_gender,x=~year,y=~suicide_ratio,type ='bar',color=suicide_gender$sex,colors=c('Paired')) %>%
  layout(title='World suicide rate per 100K by gender',
         yaxis=list(title='Suicide rate per 100K'))
```
### Overview of world suicide rate and founding
* Downward trend from 1995 to 2015

* Age group of people over 75 years old is the group which committed suicide the most over 20 years.

* Top 10 regions with the highest suicide rate in 2015: Lithuania, South Korea, Slovenia, Latvia, Hungary, Uruguay, Japan, Ukraine, Russia, Croatia.

* More male die from suicide than women.

### Interpretation and reasonings
##### Why more male die from suicide than women?
*This is also known as the gender paradox in suicide.

* Suicide attempts are between two and four times more frequent among females.

* Researchers have partly attributed the difference between attempted and completed suicides among the sexes to males using more lethal means to end their lives.

##### Why most of countries in top10 suicide rate are in East Europe?
* Economic crisis after collapse of USSR.
* Civil war.

### Suicide prevention
- September 10 has been observed as 'World Suicide Prevention Day' in partnership with the International Association for Suicide Prevention and the World Health Organization.

- South Korea Suicide Hotlines
http://www.suicide.org/hotlines/international/south-korea-suicide-hotlines.html

- Talk therapy

- Suicide Survivors are Heroes



